\chapter*{Acknowledgements}

\addcontentsline{toc}{chapter}{Acknowledgements}

This dissertation has been produced with guidance and support from many people, without which it would not have seen the light of day.

Primarily, I must thank my supervisor Andy Pilkington. His guiding hand and experience over the course of this project has at times been the only thing keeping it on the straight and narrow. I am indebted to his help and his willingness to put up with me over the course of this year, and this work is built around the backbone he provided.

Secondly, thanks to the Manchester Particle Physics group, for providing a terrific environment for work, for all of the guidance and for all of the socialising. Particular mention must go to my colleagues on the ATLAS project; Jacob, Jonathan
, Yaadav and Agni for all of their invaluable experience and assistance throughout the year. Additionally, I must thank Sabah for providing me with his help and knowledge without hesitation when I needed it.

From my time at undergraduate, I must thank all my friends, who while provided encouragement and support from all the myriad of locations around the world they've run off to, and I must thank my undergraduate Director of Studies, Dave Green, for his advice during my time under his watch and for permitting me to use his \LaTeX template.

Finally however, none of this would have occurred were it not for the support of my dad, mum and brother. Their unwavering and tireless support, endless encouragement and motivating prods and willing to put up with whatever inconvenience I could give them carried this work through to its close.

To all those mentioned here, I owe you the most sincere gratitude. Thank you for everything.
\cleardoublepage

\endinput
